---
title: 'Effect of digestion on composition and more'
output: pdf_document
author: Sasha D. Hafner
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

# Vars 
```{r}
dw$dpH <- dw$pH.dig - dw$pH.ref
```

# Plots
```{r}
ggplot(dat, aes(source, pH, colour = relDiff.frac)) +
  geom_point() +
  theme(legend.position = 'top', axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

ggplot(dw, aes(pH.ref, pH.dig)) +
  geom_point()

ggplot(dw, aes(pH.ref, dpH)) +
  geom_point()
```

# Stats
```{r}
m1 <- lm(pH.dig ~ pH.ref, data = dw)
summary(m1)
```

Try robust regression.

```{r}
m2 <- MASS::rlm(pH.dig ~ pH.ref, data = dw)
summary(m2)
```

Digestate pH does seem correlated with raw pH but only with robust regression.
Issue seems to decrease in change in pH at higher raw pH.

```{r}
t.test(dw$dpH)
```

Clearly pH does increase, according to a one-sample t-test.

But look at change in pH:

```{r}
m3 <- lm(dpH ~ pH.ref, data = dw)
summary(m3)
```

```{r}
m4 <- MASS::rlm(dpH ~ pH.ref, data = dw)
summary(m4)
```

Interestingly, the change in pH is negatively correlated with initial pH.

Change DM change.

```{r}
dw$dDM <- dw$DM.dig - dw$DM.ref
t.test(dw$dDM)
```

And DM clearly decreases.



