---
title: 'Stats 1: Effect of digestion on composition and more'
output: pdf_document
author: Sasha D. Hafner
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

# Vars 
Difference in pH due to digestion.
```{r}
dw$dpH <- dw$pH.dig - dw$pH.ref
dw$dDM <- dw$DM.dig - dw$DM.ref
```

```{r}
table(dw$source)
```

# Plots
```{r}
ggplot(dat, aes(source, pH, colour = relDiff.frac)) +
  geom_point() +
  theme(legend.position = 'top', axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

ggplot(dw, aes(pH.ref, pH.dig, colour = slurry.major.ref)) +
  geom_point()

ggplot(dw, aes(pH.ref, dpH, colour = slurry.major.ref)) +
  geom_point()
```

# Stats
```{r}
m1 <- lm(pH.dig ~ pH.ref + slurry.major.ref, data = dw)
summary(m1)
```

```{r}
m2 <- lm(pH.dig ~ pH.ref, data = dw)
summary(m2)
```
Try robust regression.

```{r}
m3 <- MASS::rlm(pH.dig ~ pH.ref + slurry.major.ref, data = dw)
summary(m3)
```

```{r}
m4 <- MASS::rlm(pH.dig ~ pH.ref, data = dw)
summary(m4)
```

Digestate pH does seem correlated with raw pH but only with robust regression.
Issue seems to be a decrease in change in pH at higher raw pH.
So post digestion pH seems to be the same regardless of raw pH.
Seems plausible.
Say low raw pH is caused by a lot of VFAs, which then have no effect on digestate pH.

Simpler question: how does digestion change pH and DM?

```{r}
t.test(dw$dpH)
```

Clearly pH does increase, according to a one-sample t-test.
Mixed-effects model more appropriate.

```{r}
m5pH <- lmer(dpH ~ (1|source), data = dw)
summary(m5pH)
confint(m5pH)
```

Compare among animal types.

```{r}
m6pH <- lmer(dpH ~ slurry.major.ref +  (1|source), data = dw)
summary(m6pH)
```

Not clearly smaller for pig.

But we don't care about standard deviation estimates, so simpler to explain mean by study.

```{r}
dws <- aggregate2(dw, c('dpH', 'dDM'), by = 'source', FUN = list(mean = mean, n = function(x) sum(!is.na(x))))
```

```{r}
t.test(dws$dpH.mean)
```

And repeat for DM:

```{r}
t.test(dw$dDM)
```

Clearly DM does increase, according to a one-sample t-test.
Mixed-effects model more appropriate.

```{r}
m5DM <- lmer(dDM ~ (1|source), data = dw)
summary(m5DM)
confint(m5DM)
```

Compare among animal types.

```{r}
m6DM <- lmer(dDM ~ slurry.major.ref +  (1|source), data = dw)
summary(m6DM)
```

Some evidence change is smaller for pigs.

We don't care about standard deviation estimates, so simpler to explain mean by study.

```{r}
t.test(dws$dDM.mean)
```

