---
title: 'Stats to parse out DM and pH effects'
output: pdf_document
author: Sasha D. Hafner
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

# Plots

```{r}
dat$DMgrp <- cut(dat$DM, 4)
dat$pHgrp <- cut(dat$pH, 4)
dd <- subset(dat, !is.na(pH) & !is.na(DM))
```

```{r}
ggplot(dd, aes(pH, emis.perc, colour = source)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ DMgrp, ncol = 4)
```

```{r}
ggplot(dd, aes(DM, emis.perc, colour = source)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ pHgrp, ncol = 4)
```

Not pretty but there do seem to be correlations for both DM and pH.

How did digestion change DM and pH?

```{r}
dd <- subset(dd, !is.na(relDiff.set))
ggplot(dd, aes(DM, pH, colour = relDiff.frac, group = interaction(source, relDiff.set))) +
  geom_line(colour = 'gray85') +
  geom_point() +
  theme_bw()
```

Generally increased pH and decreased DM, as expected in the absence of added high DM substrates.

# Models

Summary.

```{r}
dfsumm(dat[, c('source', 'app.meth', 'DM', 'pH', 'relDiff.frac', 'emis.perc')])
```

Overall digestion effect (not expecting much)

```{r}
m1 <- lm(emis.perc ~ source + app.meth + relDiff.frac, data = dat)
summary.aov(m1)
tail(coef(m1))
```

Nothing.

Separate DM and pH effects.

```{r}
m2 <- lm(emis.perc ~ pH + DM + relDiff.frac + source + app.meth, data = dat)
summary.aov(m2)
drop1(m2, test = 'F')
```

```{r}
m3 <- lm(emis.perc ~ pH + DM + source + app.meth, data = dat)
summary.aov(m3)
drop1(m3, test = 'F')
cc <- coef(m3)[1:3]
cc
confint(m3)[1:3, ]
cc['pH'] / cc['DM']
```

m3 is where it gets interesting.
Clear positive effects of DM and pH, although the pH effect seems small (+4% emission (frac. of TAN) per unit pH change).
Effect of pH is about 2x DM effect (1 unit pH change : 1% DM change).

```{r}
plot(m3, ask = FALSE)
```

We might look at relative effect.

```{r}
m4 <- lm(log10(emis.perc) ~ pH + DM + source + app.meth, data = dat)
summary.aov(m4)
cc <- 100 * (10^coef(m4)[1:3] - 1)
cc
100 * (10^confint(m4)[1:3, ] - 1)
cc['pH'] / cc['DM']
```

38% increase in emission with a 1 unit increase in pH, seems plausible.
Effect of pH here 3x DM effect. 
Seems more plausible.

```{r}
plot(m4, ask = FALSE)
```

# Mixed-effects model
As Johanna pointed out, `source` is really a random effect.

```{r}
dfsumm(dd[, c('pH', 'DM', 'app.meth', 'source')])
dsub <- subset(dat, !is.na(pH) & !is.na(DM) & !is.na(app.meth))
m5 <- lmer(log10(emis.perc) ~ pH + DM + app.meth + (1|source), data = dsub)
```

```{r}
summary(m5)
ccm <- 100 * (10^fixef(m5) - 1)
ccm
100 * (10^confint(m5)[4:5, ] - 1)
ccm['pH'] / ccm['DM']
```

Large *t* statistics and similar coefficients as classical linear model `m4`.
Compare:

```{r}
cc
ccm
```

But, let's try likelihood ratio test (I think that's what `anova()` is doing here).

```{r}
m0 <- lmer(log10(emis.perc) ~ app.meth + (1|source), data = dsub)
m0pH <- lmer(log10(emis.perc) ~ DM + app.meth + (1|source), data = dsub)
m0DM <- lmer(log10(emis.perc) ~ pH + app.meth + (1|source), data = dsub)
anova(m5, m0pH)
anova(m5, m0DM)
anova(m5, m0)
```

Look at raw and digestate separately.

```{r}
m4r <- lm(log10(emis.perc) ~ pH + DM + app.meth, data = dsub, subset = relDiff.frac == 'reference')
summary(m4r)
drop1(m4r, test = 'F')
```

```{r}
m4d <- lm(log10(emis.perc) ~ pH + DM + app.meth, data = dsub, subset = relDiff.frac == 'digestate')
summary(m4d)
drop1(m4d, test = 'F')
```

# Predictions

Might be interesting to see how predictions here stack up against current DK EFs (based on ALFAM2).

```{r}
dat$si <- factor(as.integer(factor(dat$source)))
m6 <- lm(log10(emis.perc) ~ pH + DM + si + app.meth, data = dat)
coef(m6)
preddat <- data.frame(scenario = c('ref', 'dig', 'dig+'), si = factor(19), 
                      app.meth = 'trailing hose',
                      DM = c(6.5, 5.1, 6.5), pH = c(7, 7.9, 7.9))
preddat$emis <- 10^predict(m6, newdata = preddat)
```

```{r}
knitr::kable(preddat)
```

# Different approach

```{r}
ddw <- subset(dw, !is.na(emis.perc.ref + emis.perc.dig + DM.ref + DM.dig + pH.ref + pH.dig))
ddw$demis <- ddw$emis.perc.dig - ddw$emis.perc.ref
ddw$dpH <- ddw$pH.dig - ddw$pH.ref
ddw$dDM <- ddw$DM.dig - ddw$DM.ref
```

```{r}
m7 <- lm(log10(emis.perc.dig) ~ log10(emis.perc.ref) + dpH + dDM, data = ddw)
summary(m7)
```

```{r}
m8 <- lm(demis ~ dpH + dDM, data = ddw)
summary(m8)
```

Wow.
Nothing going on.

```{r}
hist(ddw$dDM, breaks = 20)
```

```{r}
hist(ddw$dpH, breaks = 20)
```

Include interaction.

```{r}
m9 <- lm(log10(emis.perc.dig) ~ log10(emis.perc.ref) + dpH*dDM, data = ddw)
summary(m9)
```

