---
title: 'Predictions of pH change due to anaerobic digestion'
output: pdf_document
author: Sasha D. Hafner
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r}
source('NH3_mods.R')
```

Use eqSpec to calculate equilibrium speciation.

# 1. Raw pig slurry
Define composition.
```{r}
totp1 <- c(H. = 0.0, NH3 = 0.208, H2CO3 = 0.112, K. = 0.0, Na. = 0.0, Cl. = 0.0, HAc = 0.21)
```

Calculate pH.
```{r}
eqp1 <- eqSpec(tot = totp1, temp.c = 18, of = 'all')
-eqp1$l.a[1]
```

pH is 5.6, much lower than obs.
Adjust with KOH.

```{r}
eqp2 <- eqSpec(tot = totp1, temp.c = 18, pH = 6.87, adjpH = 'KOH', of = 'all')
totp2 <- eqp2$tot
totp2
```

That's a lot of KOH.

Check CO2 partial pressure for fun.

```{r}
eqp2$p.CO2
```

Try reducing VFA instead.

```{r}
eqp3 <- eqSpec(tot = totp1, temp.c = 18, pH = 6.87, adjpH = 'HAc', of = 'all')
totp3 <- eqp3$tot
totp3
```

That's plausible.

# 2. Raw cattle slurry
Define composition.
```{r}
totc1 <- c(H. = 0.0, NH3 = 0.149, H2CO3 = 0.129, K. = 0.0, Na. = 0.0, Cl. = 0.0, HAc = 0.193)
```

Calculate pH.
```{r}
eqc1 <- eqSpec(tot = totc1, temp.c = 18, of = 'all')
-eqc1$l.a[1]
```

pH is 5.1, even lower than obs was for pig.
Adjust with KOH.

```{r}
eqc2 <- eqSpec(tot = totc1, temp.c = 18, pH = 6.82, adjpH = 'KOH', of = 'all')
totc2 <- eqc2$tot
totc2
```

That's a ton of KOH.
Try reducing VFA instead.

```{r}
eqc3 <- eqSpec(tot = totc1, temp.c = 18, pH = 6.82, adjpH = 'HAc', of = 'all')
totc3 <- eqc3$tot
totc3
```

That's plausible.

# 3. Digestate
Define composition.
1/10th as much TIC.
```{r}
totd1 <- c(H. = 0.0, NH3 = 0.297, H2CO3 = 0.191, K. = 0.0, Na. = 0.0, Cl. = 0.0, HAc = 0.019)
```

Calculate pH.
```{r}
eqd1 <- eqSpec(tot = totd1, temp.c = 18, of = 'all')
-eqd1$l.a[1]
```

pH 9.1.
Clearly these differences in composition are *more than* enough to explain the differences in pH.

# 4. Removing VFA from raw slurry
What effect does removing VFA have?
First pig.

```{r}
totp4 <- totp3
totp4['HAc'] <- 0
totp4
```

```{r}
eqr1 <- eqSpec(tot = totp4, temp.c = 18, of = 'all')
-eqr1$l.a[1]
```

Plenty.

Then cattle.

```{r}
totc4 <- totc3
totc4['HAc'] <- 0
totc4
```

```{r}
eqr2 <- eqSpec(tot = totc4, temp.c = 18, of = 'all')
-eqr2$l.a[1]
```

Plenty of effect here too.

# Conclusions
1. We cannot completely explain observed pH values of raw slurry or digestate, but this isn't too worrying because we don't have complete information on composition.
2. Observed differences in VFA and TAN in raw vs. digested slurry are more than enough to explain the observed pH difference.
3. Simply removing raw slurry VFA is enough to explain the observed pH differences.



